---
title: "Seurat Pipeline"
author: "Ferdinand"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries


```{r}
library(Seurat)
library(tidyverse)
library(hdf5r)
```

## Load NSCLC dataset


```{r }
nsclc.data = Read10X_h5(filename = "C:/Users/F/Nextcloud/TEst2/Courses/Bioinformagician/Seurat Pipeline/Data/20k_NSCLC_DTC_3p_nextgem_intron_Multiplex_count_raw_feature_bc_matrix.h5")

###what are the modalities and how is the right matrix called?
str(nsclc.data)

cts = nsclc.data$`Gene Expression`

```

##Create Seurat object with the non-normalized data
#min.cells = 3 only genes that are expressed in at least 3 cells
#min.features = 200 only cells that express at least 200 genes (aka. features)

```{r}
nsclc.sobject = CreateSeuratObject(counts = cts, project = "NSCLC", min.cells = 3, min.features = 200)
#str(nsclc.sobject)
nsclc.sobject
```

##Quality Control
#1. Percentage of mitochondrial genes

```{r}
#saves in % in another variable in the metadata of the Seurat object
nsclc.sobject[["perc.mt"]] = PercentageFeatureSet(nsclc.sobject, pattern = "^MT-")

View(nsclc.sobject@meta.data)

VlnPlot(nsclc.sobject, features = c("nFeature_RNA", "nCount_RNA", "perc.mt"), ncol =3)
FeatureScatter(nsclc.sobject, feature1 = "nCount_RNA", feature2 = "nFeature_RNA") + geom_smooth(method = "lm")
```

# 2. Filtering
```{r}
nsclc.sobject <- subset(nsclc.sobject, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & perc.mt < 5)


```

# 3. Normalize data

```{r}
nsclc.sobject <- NormalizeData(nsclc.sobject,  normalization.method = "LogNormalize", scale.factor = 10000)
str(nsclc.sobject)
```

# 4. Identify highly variable features
```{r}
nsclc.sobject <- FindVariableFeatures(nsclc.sobject, selection.method = "vst", nfeatures = 2000)

# Identify the 50 most highly variable genes
top50 <- head(VariableFeatures(nsclc.sobject), 50)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(nsclc.sobject)
LabelPoints(plot = plot1, points = top50, repel = TRUE)
```
# 5. Scaling 

```{r}
all.genes <- rownames(nsclc.sobject)
nsclc.sobject <- ScaleData(nsclc.sobject, features = all.genes)

str(nsclc.sobject)
```

# 6. Linear dimensionality reduction

```{r}
nsclc.sobject <- RunPCA(nsclc.sobject, features = VariableFeatures(object = nsclc.sobject))

```

# visualize PCA results
```{r}
print(nsclc.sobject[["pca"]], dims = 1:5, nfeatures = 5)
DimHeatmap(nsclc.sobject, dims = 1, cells = 500, balanced = TRUE)
```


# determine dimensionality of the data
```{r}
ElbowPlot(nsclc.sobject)
```

# 7. Clustering
```{r}
nsclc.sobject <- FindNeighbors(nsclc.sobject, dims = 1:15)
```

# understanding resolution

```{r}
nsclc.sobject <- FindClusters(nsclc.sobject, resolution = c(0.1,0.3, 0.5, 0.7, 1))
View(nsclc.sobject@meta.data)

DimPlot(nsclc.sobject, group.by = "RNA_snn_res.0.5", label = TRUE)

```

# setting identity of clusters
```{r}
Idents(nsclc.sobject)
Idents(nsclc.sobject) <- "RNA_snn_res.0.1"
Idents(nsclc.sobject)
```

# non-linear dimensionality reduction
```{r}
reticulate::py_install(packages = 'umap-learn')

nsclc.sobject <- RunUMAP(nsclc.sobject, dims = 1:15)
```

# note that you can set `label = TRUE` or use the LabelClusters function to help label individual clusters

```{r}
DimPlot(nsclc.sobject, reduction = "umap")

```
